---
title: "MicrobiomeDada2"
author: "Reimon Fernandez"
date: "2024-04-16"
output: pdf_document
---

# Load required packages
```{r}
library(dada2)
```

# Load sequences
```{r}
path="sequences"
list.files(path)
```
# read in file names
```{r}
# Forward filenames have format: SAMPLENAME_R1_001.fastq 
fnFs <- sort(list.files(path, pattern="_R1_001.fastq", full.names = TRUE))
# Extract sample names, assuming filenames have format: SAMPLENAME_XXX.fastq
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
```

#to see the quality plot
```{r}
plotQualityProfile(fnFs[1:2])
```

# filter and trim
```{r}
# Place filtered files in filtered/ subdirectory
filtFs <- file.path(path, "filtered", paste0(sample.names, "_F_filt.fastq.gz"))
names(filtFs) <- sample.names

# quality decreases @ 120
out <- filterAndTrim(fnFs, filtFs, truncLen=c(120),
              maxN=0, maxEE=c(2), truncQ=2, rm.phix=TRUE,
              compress=TRUE, multithread=FALSE) # On Windows set multithread=FALSE
head(out)
```
# error rate calc AI 
```{r}
errF <- learnErrors(filtFs, multithread=FALSE)
plotErrors(errF, nominalQ=FALSE)
```
# Sample inference (id number of unique sequences)
```{r}
dadaFs <- dada(filtFs, err=errF, multithread=FALSE)
```

#construst sequence table 
```{r}
seqtab <- makeSequenceTable(dadaFs)
dim(seqtab)
#there are 34 samples with 819 unique sequences

# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
```


#remove chimeras; remove nything that is not a single species 
```{r}
seqtab.nochim <- removeBimeraDenovo(seqtab, method="consensus", multithread=FALSE, verbose=TRUE)
dim(seqtab.nochim)
```

# track reads through a pipeline
```{r}
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), rowSums(seqtab.nochim))
colnames(track) <- c("input", "filtered", "denoisedF", "nonchim")
rownames(track) <- sample.names
head(track)
```

# save seqtab.nochim as an R file

```{r}
save(seqtab.nochim,file="RData/seqtab.nochim.RData")
```

# load seqtab.nochim to start here
```{r}
load("RData/seqtab.nochim.RData")
```
# Assign taxonomy
```{r}
taxa=assignTaxonomy(seqtab.nochim,"silva_nr99_v138.1_wSpecies_train_set.fa.gz")
#download the silva species database from https://zenodo.org/records/4587955
```

#save taxonomy as a file
```{r}
save(taxa,file="RData/taxa.RData")
```